---
title: "homework2"
author: "hurricane"
date: "10/5/2021"
output: github_document
---

```{r,echo=FALSE,message=FALSE}
library(tidyverse)
library(readxl)
```

# problem 1

## Read and clean Mr.Trash Wheel sheet

```{r}
trash_wheel=
  read_excel("data/trash_wheel_collection.xlsx",range="A2:N408") %>%   
  janitor::clean_names() %>% 
  na.omit(dumpster) %>% 
  mutate(sports_balls=round(sports_balls,0)) 
```



## Read and clean precipatation data for 2018 and 2019
```{r}
precip2018_df=
  read_excel("data/trash_wheel_collection.xlsx",sheet=5,range="A2:B15") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate(year="2018") %>% 
  select(year,everything())


precip2019_df=
  read_excel("data/trash_wheel_collection.xlsx",sheet=4,range="A2:B15") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate(year="2019") %>% 
  select(year,everything())


precip_2018_2019_df=bind_rows(precip2018_df,precip2019_df) %>% 
  janitor::clean_names() %>% 
  mutate(month.name[month]) %>% 
  select(!month) %>% 
  rename("month"="month.name[month]") %>% 
  select(year,month,total) 


```

The first sheet of Mr.Trash Wheel dataset includes `r nrow(trash_wheel)` rows and `r ncol(trash_wheel)` variables which provide us with information about weight of trash collected by Mr.Trash Wheel from May 2014 to June 2019. Furthermore, it lists weight of different types of trash which shows in `r colnames(trash_wheel[7:8])` columns and so on.

Mr.Trash Wheel dataset also includes every month's precipitation data for 2018 and 2019, which has `r nrow(precip_2018_2019_df) ` rows and `r ncol(precip_2018_2019_df)` columns.


The total precipitation in 2018 is `r sum(precip_2018_2019_df[which(precip_2018_2019_df[,1]==2018),3])` and the median number of sports balls in a dumpster in 2019 is `r median(trash_wheel$sports_balls)`.









# Problem 2

### Import and clean pols-month data
```{r,message=FALSE}
pols_month_df=
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon,into=c("year","month","day"),sep="-") %>%
  mutate(year=as.integer(year),month=as.integer(month),day=as.integer(day)) %>% 
  mutate(month.name[month]) %>%
  select(-month) %>% 
  rename("month"="month.name[month]") %>% 
  mutate(president=ifelse(prez_gop==1,"gop","dem")) %>%
  select(!prez_gop & !prez_dem & !day) %>% 
  mutate(month=substring(month,0,3),year=as.numeric(year)) %>% 
  select(year,month,everything()) %>% 
  mutate(year=as.integer(year))
```
  
### Import and clean snp data
```{r,message=FALSE}
snp_df=
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  mutate(date=lubridate::mdy(date)) %>% 
  separate(date,into=c("year","month","day"),sep="-") %>% 
  mutate(year=as.integer(year),month=as.integer(month),day=as.integer(day)) %>% 
  mutate(month.name[month])%>% 
  select(-month) %>% 
  rename("month"="month.name[month]") %>% 
  mutate(month=substring(month,0,3)) %>% 
  select(year,month,everything()) %>% 
  mutate(year=ifelse(year>2049,year-100,year)) %>% 
  select(year,everything(),-day) %>%
  arrange(year,month)

```


### Import and clean unemplyment data
```{r,message=FALSE}
unemployment_df=
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(2:13, names_to="month",values_to="unemployment_rate"
  ) %>% 
  janitor::clean_names() %>% 
  arrange(year,month) %>% 
  mutate(year=as.integer(year))
```

### Join three dataset together
```{r}
merge_snp_unemployment_df1=
  left_join(pols_month_df,snp_df,by=c("year","month"))
  
merge_snp_unemployment_df2=
  left_join(merge_snp_unemployment_df1,unemployment_df,by=c("year","month"))
```

After cleaning, the file "pols-month"  contains `r nrow(pols_month_df)` observations of `r ncol(pols_month_df)` variables related to the number of national politicians who are democratic or republican from 1947 to 2015. 

After cleaning, the file “snp” contains `r nrow(snp_df)` observations of `r ncol(snp_df)` variables related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole.

After cleaning, the file "unemployment" contains `r nrow(unemployment_df)` observations of `r ncol(unemployment_df)` variables.

The three-joint dataset contains `r nrow(merge_snp_unemployment_df2)` observations of `r ncol(merge_snp_unemployment_df2)` variables from 1947 to 2015. The key variables include president, unemployment rate and so on.


## Problem 3

### Clean baby name
```{r,message=FALSE}
baby_name_df=
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name=tolower(childs_first_name)) %>% 
  mutate(childs_first_name=str_to_title(childs_first_name)) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="ASIAN AND PACI","ASIAN AND PACIFIC ISLANDER")) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="BLACK NON HISP","BLACK NON HISPANIC")) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="WHITE NON HISP","WHITE NON HISPANIC")) %>% 
  distinct() %>% 
  arrange(year_of_birth)
```

### Well-structured, reader-friendly table 
```{r,message=FALSE}
baby_name_df=
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name=tolower(childs_first_name)) %>% 
  mutate(childs_first_name=str_to_title(childs_first_name)) %>% 
  distinct() %>% 
  arrange(ethnicity,year_of_birth) %>% 
  filter(childs_first_name=="Olivia") %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="ASIAN AND PACI","ASIAN AND PACIFIC ISLANDER")) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="BLACK NON HISP","BLACK NON HISPANIC")) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="WHITE NON HISP","WHITE NON HISPANIC")) %>% 
  select(-count,-gender,-childs_first_name) %>% 
pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
    ) 
```


## a table showing the most popular name among male children over time
```{r,message=FALSE}
baby_name_df=
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name=tolower(childs_first_name)) %>% 
  mutate(childs_first_name=str_to_title(childs_first_name)) %>% 
  distinct() %>% 
  arrange(ethnicity,year_of_birth) %>% 
  filter(gender=="MALE") %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="ASIAN AND PACI","ASIAN AND PACIFIC ISLANDER")) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="BLACK NON HISP","BLACK NON HISPANIC")) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="WHITE NON HISP","WHITE NON HISPANIC")) %>% 
  arrange(year_of_birth) %>% 
  filter(rank=="1") %>% 
  select(-gender,-count,-rank) %>% 
  pivot_wider(
    names_from="year_of_birth",
    values_from="childs_first_name"
  ) 
```

## A scatter plot
```{r}
baby_name_df=
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name=tolower(childs_first_name)) %>% 
  mutate(childs_first_name=str_to_title(childs_first_name)) %>% 
  distinct() %>% 
  arrange(ethnicity,year_of_birth) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="ASIAN AND PACI","ASIAN AND PACIFIC ISLANDER")) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="BLACK NON HISP","BLACK NON HISPANIC")) %>% 
  mutate(ethnicity = replace(ethnicity,ethnicity=="WHITE NON HISP","WHITE NON HISPANIC")) %>% 
  arrange(year_of_birth) %>% 
  filter(gender=="MALE",ethnicity=="WHITE NON HISPANIC",year_of_birth=="2016") %>%
  select(childs_first_name,count,rank)
#library(ggplot2)

ggplot(baby_name_df,aes(x=rank,y=count),main="Relation of Rank and count of Name  Scatter Plot") + 
  geom_point(size=0.5) +
  geom_rug(col="steelblue",alpha=0.1, size=1.5) + ggtitle("Relation of Rank and count of Name  Scatter Plot")



```

